---
title: "Numeric EDA"
output: github_document
date: "2022-10-06"
---

```{r setup, include = FALSE}
library(tidyverse)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load the weather data

```{r data}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

## Using `group_by`

Datasets are often comprised of groups defined by one or more (categorical) variable; `group_by()` makes these groupings explicit so that they can be included in subsequent operations. For example, we might group weather_df by name and month:

```{r}
weather_df %>%
  group_by(name, month) # Doesn't change dataset itself, but makes a new layer in the df
```
Several important functions respect grouping structures. You will frequently use `summarize` to create one-number summaries within each group, or use `mutate` to define variables within groups. 

Because these (and other) functions will use grouping information if it exists, it is sometimes necessary to remove groups using `ungroup()`.

## `summarize()`

Summarizing gives you group-specific summaries. The `n()` function from `tidyverse` lets you count things in a group.

The result is a dataframe that includes the grouping variable(s) and the desired summary.

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarise(
    n_obs = n()
  )

weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    n_obs = n()
  )
```

Can we count in other ways?

To count things, you could use `count()` in place of `group_by()` and `summarize()` if you remember that this function exists. 

```{r}
weather_df %>% 
  count(name, month)
```

Don't use `table()`!!

`count()` is a useful tidyverse alternative to Base R’s `table` function. Both functions produce summaries of how often values appear, but `table`’s output is of class table and is hard to do any additional work with, while `count` produces a dataframe you can use or manipulate directly. 

```{r, eval = FALSE}
weather_df %>%
  pull(month) %>% 
  table
```

You can have multiple summaries. `n_distinct()` counts up the number of distinct observations in a dataset, or for a specific column. As an example, we count the number of observations in each month and the number of distinct values of date in each month.

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarise(
    n_obs = n(), 
    n_dist = n_distinct(date)
  )
```

## A brief aside on 2x2 Tables

Nothing in our dataset is conducive to a 2x2 structure, but we can try...
In a contrived example, let’s say you want to look at the number of cold and not-cold days in Central Park and Waterhole. 

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```



