---
title: "Visualization"
subtitle: "Part I"
output: github_document
date: "2022-09-29"
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
```

## Load the weather data

```{r, warning=FALSE, message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots!!

Create my first scatterplot ever. 
We first specify the data, then the aesthetic mappings. 

```{r scatterplot}
ggplot(weather_df, aes(x = tmin, y = tmax)) + #ggplot sets up axis ranges based on aes()
  geom_point()
```

New approach (using pipes), same plot. Using pipes, we can mutate, filter, select, before making our plot.

```{r pipeplot}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Save and edit plot object (Less common workflow). You can create a ggplot object, and then add `geom` functions, print or save later, etc.

```{r plotobject}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot...

Start with the same one and make it fancy! We are going to assign the `color` of the points to the `name` of the weather station using `aes`. We will also add another geometry, `geom_smooth`. 

```{r advanced}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

What about the `aes` placement? When we move aesthetics inside specific geometries, the color applies only to the scatterplot, not the smooth line plot! If you put aesthetics in ggplot, it is applied universally.

```{r aes, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + # Move color aes into the specific geometry (points)
  geom_smooth()
```

Let's facet some things!! (And adjust transparency)

```{r facet, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5, size = 0.3) + #Apply 50% transparency to geom_point
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) # "Nothing" defines rows (don't make rows), name defines columns

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) + #Apply transparency according to tmin
  geom_point(size = 0.3) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) # "Nothing" defines rows (don't make rows), name defines columns
```

Let's combine some elements and try a new plot.

```{r, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

