---
title: "Visualization"
subtitle: "Part II"
output: github_document
date: "2022-10-04"
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load the weather data

```{r, warning=FALSE, message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```
As a starting point, let’s revisit the scatterplot of `tmax` against `tmin` made in Visualization Pt 1.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

## Labels

We can add labels using the `labs()` function. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package",
    color = "Location"
  )
```

## Scales

You can adjust the scales and add labels to the scales. We can also transform the scales using `trans()` (e.g. square root, etc.)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package",
    color = "Location"
  ) +
  scale_x_continuous(breaks = c(-10, 0, 15),
                     labels = c("-10C", "0", "15C")) +
  scale_y_continuous(
    trans = "sqrt"
  )
```

## Colours

We can manually adjust colours using `scale_colour_hue()`, or use the `viridis` package. In general, encourage using the viridis colour palette.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package",
    color = "Location"
  ) +
  scale_colour_hue(
    name = "Location",
    h = c(100, 300))

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package",
    color = "Location"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE #Makes it a discrete rather than continuous colour palette
  )
```

## Themes

```{r}
ggp_weather = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_weather
```

Move the legend to the bottom of the plot. We can get adjust the grey background, grid bars, etc.

```{r}
ggp_weather + 
  theme_minimal() +
  theme(legend.position = "bottom")
```

Make sure your theme functions in the correct order! Applying an overall theme overrides individual adjustments to the plot element. Here, we can see the legend is no longer at the bottom.

```{r}
ggp_weather + 
  theme(legend.position = "bottom") +
  theme_minimal()
```

## Setting options

You can set global options to control all ggplots you make in a R markdown. (see P8105 website)

## Data argument in `geom_()`

We want to use different geometries for Central Park and Waikiki data. 

```{r}
central_park_df = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki_df = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(waikiki_df, aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_line(data = central_park_df)

```

## Patchwork...

We want to make a panel of plots with the same dataset (same aesthetic mappings, etc.) easily using `facet`. However, we might want to have a panel of fundamentally different figures using `patchwork`.

```{r}
tmax_tmin_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, colour = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_density_plot = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) + 
  theme(legend.position = "none")

seasonality_plot = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, colour = name)) + 
  geom_point(alpha = .5) + 
  geom_smooth(se = FALSE)
  theme(legend.position = "none")

(tmax_tmin_plot + prcp_density_plot)/ seasonality_plot # This requires loading the patchwork library!
```

## Data manipulation





