---
title: "Visualization"
subtitle: "Part I"
output: github_document
date: "2022-09-29"
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
```

## Load the weather data

```{r, warning=FALSE, message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots!!

Create my first scatterplot ever. 
We first specify the data, then the aesthetic mappings. 

```{r scatterplot}
ggplot(weather_df, aes(x = tmin, y = tmax)) + #ggplot sets up axis ranges based on aes()
  geom_point()
```

New approach (using pipes), same plot. Using pipes, we can mutate, filter, select, before making our plot.

```{r pipeplot}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Save and edit plot object (Less common workflow). You can create a ggplot object, and then add `geom` functions, print or save later, etc.

```{r plotobject}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot...

Start with the same one and make it fancy! We are going to assign the `color` of the points to the `name` of the weather station using `aes`. We will also add another geometry, `geom_smooth`. 

```{r advanced}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

What about the `aes` placement? When we move aesthetics inside specific geometries, the color applies only to the scatterplot, not the smooth line plot! If you put aesthetics in ggplot, it is applied universally.

```{r aes, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + # Move color aes into the specific geometry (points)
  geom_smooth()
```

Let's facet some things!! (And adjust transparency)

```{r facet, warning = FALSE, message=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5, size = 0.3) + #Apply 50% transparency to geom_point
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) # "Nothing" defines rows (don't make rows), name defines columns

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) + #Apply transparency according to tmin
  geom_point(size = 0.3) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) # "Nothing" defines rows (don't make rows), name defines columns
```

Let's combine some elements and try a new plot.

```{r, warning = FALSE, message=FALSE}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

## Some small notes

How many geoms have to exist?
```{r, warning=FALSE, message=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
```

For scatterplots, you can use a neat geom! First, install the `hexbin` package.

```{r, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_density2d() + 
  geom_point(alpha = 0.3)
```

## Univariate plots

Histograms are great!!

```{r hist}
weather_df %>% 
  ggplot(aes(x = tmin)) + 
  geom_histogram()
```

Can we add colour...
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) + 
  geom_histogram()
```

It sort of works...location looks funny and only the outline is coloured.

```{r, warning=FALSE, message=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + #Use fill instead of color
  geom_histogram(position = "dodge") #use dodge
```
Usually, we avoid putting multiple variables in the same plot. Better using facet.

```{r, message=FALSE, warning=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram() + 
  facet_grid(. ~ name)
```

If we want to compare distributions across levels, maybe avoid using a histogram...

## Density

Let's try a new geometry! Density plots!

```{r, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density(alpha = 0.3, adjust = .5) #Adjust gives you more bumpy data
```

## Box plots
What about box plots??
```{r, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) + 
  geom_boxplot()
```

## Trendy plots :-)

Violin plots are like a cross between a box plot and density plot. Might be more informative as a box plot.

```{r, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin(alpha = 0.4) + 
  stat_summary(fun = "median")
```

Ridge plots -- the most popular plot of 2017 (must load `ggridges` package). Very viable for categorical variables with MANY levels (compared to a box plot or violin plot). 

```{r, warning=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) + 
  geom_density_ridges()
```

## Save and embed

Let's save a scatterplot. We should define an object for the plot. If you have a lot of plots, maybe save them in a `results` folder.

```{r save}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)
```

What about embedding...
We can play with the code chunk options.

```{r, fig.width = 12, fig.height = 6}
weather_plot
```



